---
title: "DSiR00"
author: "Robert A. Stevens"
date: "March 8, 2016"
output: html_document
---

**TODO:**

1. Finish outline

2. Add R code

3. Run R code and check

4. Do exercises

*Data Science in R: A Case Studies Approach to Computational Reasoning and Problem Solving*

by Deborah Nolan and Duncan Temple Lang

http://www.rdatasciencecases.org

## Chapter 2. Modeling Runners’ Times in the Cherry Blossom Race 

by Daniel Kaplan and Deborah Nolan

## 2.1 Introduction

Figure 2.1

### 2.1.1 Computational Topics

## 2.2 Reading Tables of Race Results into R

Figure 2.2

Figure 2.3

```{r, comment=NA}

```

```{r, comment=NA}

```
```{r, comment=NA}

```
```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

## 2.3 Data Cleaning and Reformatting Variables

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.4

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.5

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

## 2.4 Exploring the Run Time for All Male Runners

```{r, comment=NA}

```

Figure 2.6

### 2.4.1 Making Plots with Many Observations

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.7

```{r, comment=NA}

```

Figure 2.8

Figure 2.9

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

### 2.4.2 Fitting Models to Average Performance

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.10

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.11

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

### 2.4.3 Cross-Sectional Data and Covariates

Figure 2.12

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.13

```{r, comment=NA}

```

Figure 2.14

Figure 2.15


## 2.5 Constructing a Record for an Individual Runner across Years

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.16

## 2.6 Modeling the Change in Running Time for Individuals

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.17

```{r, comment=NA}

```

Figure 2.18

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.19

```{r, comment=NA}

```

```{r, comment=NA}

```

## 2.7 Scraping Race Results form the Web

Figure 2.20

Figure 2.21

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```
```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

## 2.8 Exercises

Q.1

Q.2

Q.3

Q.4

Q.5

Q.6

Q.7

Q.8

Q.9

Q.10

Q.11

Q.12

Q.13

Q.14

Q.15

Q.16

Q.17

Q.18

Q.19
