---
title: "DSiR00"
author: "Robert A. Stevens"
date: "March 8, 2016"
output: html_document
---

**TODO:**

1. Finish outline

2. Add R code

3. Run R code and check

4. Do exercises

*Data Science in R: A Case Studies Approach to Computational Reasoning and Problem Solving*

by Deborah Nolan and Duncan Temple Lang

http://www.rdatasciencecases.org

## Chapter 2. Modeling Runners’ Times in the Cherry Blossom Race 

by Daniel Kaplan and Deborah Nolan

## 2.1 Introduction

Figure 2.1: Screen Shot of Cherry Blossom Run Web site. This page contains links to each year's race results. The year 1999 is the earliest for which they provide data. Men's and women's results are listed separately.

### 2.1.1 Computational Topics

## 2.2 Reading Tables of Race Results into R

Figure 2.2: Screen Shot of the 2012 Male Results. This screenshot shows the results, in race order, for mean competing in the 2012 Cherry Blossom 10 Mile Run. Notice that both 5-mile times and net times are provided. We know that the Time column is net time because it is so indicated in the header of the table.

Figure 2.3: Screen Shot of Men's 2011 Race Results. This screenshot shows the results, in race order, for men competing in the 2011 Cherry Blossom road race. Notice that in 2011, 3 times are recorded - the time to compete the first 5 miles and the gun and net times for the full run. In contrast, the results from 2012 do not provide gun time.

```{r, comment=NA}

```

```{r, comment=NA}

```
```{r, comment=NA}

```
```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

## 2.3 Data Cleaning and Reformatting Variables

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.4: Box Plot of Age by Year. These side-by-side boxplots of age for each race year show a few problems with the data for 2003 and 2006. The runners in these years are unusually young.

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.5: Box Plot of Age by Year. These side-by-side boxplots of age for each race year show a reasonable age distribution. For example, the lower quartile for all years range between 29 and 32. The problems indentified earlier for 2003 and 2006 have been addressed.

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

## 2.4 Exploring the Run Time for All Male Runners

```{r, comment=NA}

```

Figure 2.6: Default Scatter Plot for Run Time vs. Age for Male Runners. This plot demonstrated that a simple plot of run time by age for the 70,000 male runners leads to such severe overplotting that the shape of the data is not descernible.

### 2.4.1 Making Plots with Many Observations

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.7

```{r, comment=NA}

```

Figure 2.8

Figure 2.9

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

### 2.4.2 Fitting Models to Average Performance

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.10

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.11

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

### 2.4.3 Cross-Sectional Data and Covariates

Figure 2.12

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.13

```{r, comment=NA}

```

Figure 2.14

Figure 2.15


## 2.5 Constructing a Record for an Individual Runner across Years

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.16

## 2.6 Modeling the Change in Running Time for Individuals

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.17

```{r, comment=NA}

```

Figure 2.18

```{r, comment=NA}

```

```{r, comment=NA}

```

Figure 2.19

```{r, comment=NA}

```

```{r, comment=NA}

```

## 2.7 Scraping Race Results form the Web

Figure 2.20

Figure 2.21

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```
```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

```{r, comment=NA}

```

## 2.8 Exercises

Q.1

Q.2

Q.3

Q.4

Q.5

Q.6

Q.7

Q.8

Q.9

Q.10

Q.11

Q.12

Q.13

Q.14

Q.15

Q.16

Q.17

Q.18

Q.19
